week3 - class RSelenium

### 2024-09-18 Webscraping-Chapter8-tutorial.Rmd
week3class

# Get anchor data

@@

#so we need to find WHERE the table is located in the html 'inspect element' in mozilla firefox or
# view page source' and you see that everything AFTER /td in the 'body' of the page seems to be
# the table we do need
soc_staff <- soc_staff %>%
    rvest::html_nodes("body") %>%
    xml2::xml_find_all("//td") %>%
    rvest::html_text()

soc_staff

#Assigning even and odd rows to 'fodd' and 'feven' (bc the output just lists the left and right side of the table interchangably)
fodd <- function(x) x%%2 != 0
feven <- function(x) x%%2 == 0

@@

>>>>

# Get anchor data -RU sociology

@@

#so we need to find WHERE the table is located in the html 'inspect element' in mozilla firefox or
# view page source' and you see that everything AFTER /td in the 'body' of the page seems to be
# the table we do need

# soc_staff_0 <- soc_staff %>%
#     rvest::html_nodes("body") %>%
#     xml2::xml_find_all("//td") %>% # xml2 here looks across multiple layers/ child-elements (hence the // !!! )
#     rvest::html_text()
# 
# 
# # BETTER?
# # using html_elements to ONLY USE ONE PACKAGE (RVEST)
soc_staff <- soc_staff %>%
    rvest::html_elements("body") %>%
    rvest::html_elements ("td") %>%
    rvest::html_text()
#   
#     
# view(soc_staff)
# 
# # Get whole table at once? 
# soc_staff_1 <- soc_staff %>%
#     rvest::html_elements("body") %>%
#     rvest::html_table()

#view(soc_staff_1)
#class(soc_staff_1) # list (you can turn into a data frame to use it for further data manipulatino fromhere on)

# (fshowdf) to make output look pretty (but i probably misunderstood what exactly Jochem said)

# get anchor data - Leiden University - PolSci 
```{r}

#do it for Leiden University 
leiden_polsci_staff <- read_html("https://www.universiteitleiden.nl/en/academic-staff/overview?pageNumber=1&faculty=social-and-behavioural-sciences&facultyinstitute=institute-of-political-science") #extracts the source html of a webpage

# div id content 
# section 
# ul class table list

leiden_polsci_staff_2 <- leiden_polsci_staff %>%
    rvest::html_elements("body") %>%
    rvest::html_elements("li")  %>%
    rvest::html_elements("strong") %>%
    rvest::html_text()

view(leiden_polsci_staff_2)

# sort rows into names and expertise 
```{r}
#Assigning even and odd rows to 'fodd' and 'feven' (bc the output just lists the left and right side of the table interchangably)
fodd <- function(x) x%%2 != 0
feven <- function(x) x%%2 == 0

@@

# gender from names 
```{r}
# voornamenbank Maartens insitute 
# e.g. Vero:
# https://nvb.meertens.knaw.nl/naam/is/Vero

# use the concatenate function cat 
# and make a loop that just fills in the name in question in the link

# INSERT CODE HERE

# Last names: https://cbgfamilienamen.nl/nfb/
```

# custom functions chunk for functions to run later on 
```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

#install.packages("kableExtra")
library(kableExtra)

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```
# Open alex 
```{r}
# SEE SUNBELT WEBSITE: Attributes tab 
# if you put html from open alex api into the read html function then you need to replace any spaces with a + symbol

# get JSON package so it can read the JSON object

```
# R Package openalexR (using r wrapper )
```{r}
# make JSON object easier to use

#install.packages("openalexR")
library(openalexR)

options(openalexR.mailto = "palmtreesdude@gmail.com")

df <- oa_fetch(entity= "author", search = "Jochem Tolsma")

fshowdf(df)

#second affiliation ofJochem is lost when using this wrapper 
# SOMETIMES INFORMATION CAN GET LOST 
df_papers <- oa_fetch(entity = "works", author.id = df$id)
df_papers$author[1] # selecting first item in column "author" which in this case is a full data base 
# OUtput gives us id from author AND CO-author

# To prevent doubles for authors that have common names we can use MULTIPLE FILTERS: eg. institution Radboud University and country NL and so on...

# find institution ids first to then filterforinstitutions:
oa_fetch(entity = "institutions", search = "Radboud University")
#output will show id: https://openalex.org/I145872427	 for Radboud University Nijmegen

# in code on Jochems in line 162 file title 45 publicationsOpenAlex.rmd

oa_fetch(entity= "author", search = "Jochem Tolsma", affiliations.institution.id = "https://openalex.org/I145872427")
#putput already has a lot of information on Jochem

# NEXT: find all papers Jochem wrote

oa_fetch (entity = "works", author.id = 'https://openalex.org/A5087380803')
 # id is the author id  from somewhere prior in Jochems script

```
## Adjacency matrix for 4 authors 
```{r}
packages <- c("tidyverse", "scholar", "openalexR", "rvest", "jsonlite")
fpackage.check(packages)
# make adjacency matrix for these four people

#Jochem tolsma
# Tom van de Meer
# Maurice Gesthuizen 
# Michael Savelkoul

# What data do we need? 
  # Authors, University, works, and co-authors 
  # All in one data frame

# Jochem Tolsma 
df_jochem <- oa_fetch(entity= "author", search = "Jochem Tolsma", affiliations.institution.id = "https://openalex.org/I145872427")

fshowdf(df_jochem)

# Maurice Gesthuizen
df_maurice <- oa_fetch(entity= "author", search = "Maurice Gesthuizen", affiliations.institution.id = "https://openalex.org/I145872427")

fshowdf(df_maurice)

# Michael Savelkoul
df_michael <- oa_fetch(entity= "author", search = "Michael Savelkoul", affiliations.institution.id = "https://openalex.org/I145872427")

fshowdf(df_michael)

# Tom van der Meer
df_tom <- oa_fetch(entity= "author", search = "Tom van der Meer", affiliations.institution.id = "https://openalex.org/I145872427")[1,]

fshowdf(df_tom)
#if output two id's: make decision rule
# here savest bet isto collect all 5 tom's there are bc we make a matrix
# if we want to look at citation scores, we have to probably only take the first id

#one vector with all authors 
all_authors <- rbind(df_jochem, df_maurice, df_michael, df_tom)

```
### works

```{r}
# Jochem's work 
df_works_jochem <- oa_fetch (entity = "works", author.id = "https://openalex.org/A5087380803")

# Maurice Gesthuizen
df_works_maurice <- oa_fetch (entity = "works", author.id = "https://openalex.org/A5007673492")

# Michael Savelkoul
df_works_michael <- oa_fetch (entity = "works", author.id = "https://openalex.org/A5048988743")

# Tom van der Meer
df_works_tom <- oa_fetch (entity = "works", author.id = "https://openalex.org/A5010780363")

# vector with all works
all_works <- rbind(df_works_tom, df_works_michael, df_works_maurice, df_works_jochem)

```
## make matrix 
```{r}
# ALT + MINUS make the ARROWS <- 
# STR + SHIFT + M makes the %>% 
# STR + SHIFT + C can make a comment of multiple lines you highlighted

# use column 'author' from works overview 
# 4x4 matrix
# decide for which time period 
  # 2 waves of 4 years: 2019-2023, 2014-2018 
  # select papers with publication date in this date frame 
    # publication year variable

# make selection of dates
#wave 1
works_wave1 <- all_works %>% filter(publication_year >= 2014 & publication_year <= 2018)
#wave 2
works_wave2 <- all_works %>% filter(publication_year >= 2019 & publication_year <= 2023)

# make empty 4x4 matrix
  # name each row and column with the identifiers of scholars

matrix_wave1 <- matrix(0, nrow = 4, ncol = 4)

rownames(matrix_wave1) <- c("https://openalex.org/A5087380803", "https://openalex.org/A5007673492", "https://openalex.org/A5048988743", "https://openalex.org/A5010780363")

colnames(matrix_wave1)<- c("https://openalex.org/A5087380803", "https://openalex.org/A5007673492", "https://openalex.org/A5048988743", "https://openalex.org/A5010780363")

print(matrix_wave1)

# loop through works

  # row 1: 
    # works[1,] # in the loop it would be 'i'

  # output shows we want information from column"au_id"
    works_wave1[1,]$author[[1]]$au_id

    # output: first listed is the SENDER
    ego <- works_wave1[1,]$author[[1]]$au_id[1]

    # the RECEIVERS are everyone BUT the ego aka. the first one
    alters <- works_wave1[1,]$author[[1]]$au_id[-1]

  # CHECK if ego is part of authors 
  # Filter only the author who are part of the other authors 
    # Check if author is in list 
      # t[t %in% authors]

  matrix_wave1[ego,alters] <- 1
  #works_wave1[ego,alters] <- works_wave1[ego,alters] + 1

# Create a 4x4 matrix filled with 0's
matrix_wave2 <- matrix(0, nrow = 4, ncol = 4)

rownames(matrix_wave2) <- c("https://openalex.org/A5087380803", "https://openalex.org/A5007673492", "https://openalex.org/A5048988743", "https://openalex.org/A5010780363")

colnames(matrix_wave2)<- c("https://openalex.org/A5087380803", "https://openalex.org/A5007673492", "https://openalex.org/A5048988743", "https://openalex.org/A5010780363")

print(matrix_wave2)

# old code from week 2 
# matrix_graph <- graph.adjacency(matrix_wave1)

# print(matrix_graph)
```

### 2024-09-18 week3RSelenium.Rmd
week3class

---
title: "t62al"
author: "by: Veroooo"
bibliography: references.bib

@@

<!--- 
HOW TO MAKE TOKENS:
install.packages("gitcreds") 
gitcreds::gitcreds_set() --->

<!--- TO MAKE NEW CHUNKS: ctrl + Alt + i --->

```{r, eval=FALSE}
########################### Title: Introduction to R for SNA Author: J Tolsma version: 30-10-2019

# start with clean workspace
rm(list = ls())

# install.packages I will need later here
install.packages("installr")  # you  first install packages
install.packages("foreign")
install.packages("tidyverse")
```

```{r}
require(foreign)  # used to read in spss data files
require(tidyverse)

# define workdirectory, note the double backslashes
# setwd("C:\Users\Vero\OneDrive - Radboud Universiteit\Documents\GitHub\labjournal")  # change to your own workdirectory
```
```{r Cheat sheet}
###########Cheat Sheet##########

# variable labels 
# c("primary", "secondary", "tertiary")

# Operator symbols: 
# $
# == logical operator. What are the other logical operators in R?
# [] is used to subset elements from an object (e.g. dataframe/vector/matrix)
#  NA is used in R to define missing values. It means Not Applicable.

# Functions:
# summary(), attributes(), attr(), table(), names(), head(), fix(), View()

# as.character() 
# as.numeric()

# in tidy: 
  #the piping operator 
  # %>%, 
  # the left is transported to the (first argument of the) function on the right. Thus, in the example below, you see that in the second and third call to mutate I don’t have to tell the function which dataset I am using.

# mean(), 
# rowMeans(), 
# rowSums, 
# cbind(), 
# is.na(), 
# ifelse().

# logicals FALSE and TRUE can be summed? (FALSE equals 0 and TRUE equals 1).
```

```{r}
# ignore the warnings ?read.spss note that I have saved the data files in a folder called
# 'addfiles'.

cv08 <- foreign::read.spss("tutorial_data\\Cultural_Changes_2008.sav", use.value.labels = T, to.data.frame = T)
cv10 <- foreign::read.spss("tutorial_data\\Cultural_Changes_2010.sav", use.value.labels = T, to.data.frame = T)

# normally I think setting use.value.labels=F is more convenient. Thus lets load the data again but
# now WITHOUT LABELS 
cv08_nolab <- foreign::read.spss("tutorial_data\\Cultural_Changes_2008.sav", use.value.labels = F, to.data.frame = T)
cv10_nolab <- foreign::read.spss("tutorial_data\\Cultural_Changes_2010.sav", use.value.labels = F, to.data.frame = T)

# finally, import the data using haven
cv08_haven <- haven::read_spss("tutorial_data\\Cultural_Changes_2008.sav")
cv10_haven <- haven::read_spss("tutorial_data\\Cultural_Changes_2010.sav")

str(cv08) # str = asks for structure of an object 
str(cv08_nolab)
str(cv08_haven)

# acsessing variable:
# datasetname$variablename

str(cv08$lftop)  # a factor
str(cv08_nolab$lftop)  # a numeric variable
str(cv08_haven$lftop)  # a 'dbl+lbl' this stands for doubles, or real numbers, which are labeled  
# next to the data itself, attributes are stored
attributes(cv08$lftop)
attributes(cv08_nolab$lftop)
attributes(cv08_haven$lftop)

# to access specific attributes
attr(cv08_haven$lftop, "labels")

summary(cv08$lftop)
summary(cv08_nolab$lftop)
summary(cv08_haven$lftop)
table(cv08_haven$lftop, useNA = "always")

#LOOKING AT DATA 
names(cv08_haven)
summary(cv08_haven)
head(cv08_haven)

# fix(cv08_haven) #will produce an error
# fix(cv08)
# View(cv08_haven)

----
```{r A.7.1 Define Missings}

str(cv08_haven$lftop)
summary(cv08_haven$lftop)
attr(cv08_haven$lftop, "labels")
table(cv08_haven$lftop, useNA = "always")

# copy the original variable in a new one, and attach it to the dataset. 
cv08$lftop_new <- cv08$lftop

# replace those values of our new variable
cv08$lftop_new[cv08$lftop_new == "Onbekend"] <- NA

#Check if recode works 
table(cv08$lftop_new, useNA = "always")

levels(cv08$lftop_new)

# as numeric variable not as a factor (categorical).
str(cv08$lftop_new)

cv08$agen <- as.numeric(as.character(cv08$lftop_new))  
# how clumsy. we first convert the factor to a string and then to a numeric variable.
table(cv08$agen, useNA = "always")

str(cv08$agen)

```{r A.7.2 Tidy }

# install.packages("tidyverse")
library(tidyverse)

# ‘mutate’ the original dataset by adding a variable.
cv08_haven <- mutate(cv08_haven, lftop_new = lftop)

#Replace missings.Note: 99 is the “onbekend” category.
cv08_haven$lftop_new <- na_if(cv08_haven$lftop_new, 99)

#Normally: combine these two steps into one
cv08_haven <- mutate(cv08_haven, lftop_new = na_if(lftop, 99))

#check
table(cv08_haven$lftop_new, useNA = "always")

```
----
```{r A.8 Recode variables}

# var006n = second variable lets us recode the variable education (in CV08)
#create a new variable educ3 with three levels:
#1. primary
#2. secondary
#3. tertiary

#levels
levels(cv08$var006n)

# table 
table(cv08$var006n, useNA = "always")

# lets make it a numeric variable first
cv08$educn <- as.numeric(cv08$var006n)
# check
table(cv08$educn, useNA = "always")

# start with an empty variable
cv08$educ3 <- NA
# fill category by category
cv08$educ3[cv08$educn == 2 | cv08$educn == 3] <- 1
cv08$educ3[cv08$educn > 3 & cv08$educn < 8] <- 2
cv08$educ3[cv08$educn > 7 & cv08$educn < 11] <- 3

# check
table(cv08$educ3, useNA = "always")

prop.table(table(cv08$educ3, useNA = "always"))

# now educ3 is a numeric variable, we want it as factor
cv08$educ3 <- as.factor(cv08$educ3)
table(cv08$educ3, useNA = "always")

levels(cv08$educ3) <- c("primary", "secondary", "tertiary")
table(cv08$educ3, useNA = "always")
```

```{r A.8.2 Tidy}

# dplyr::recode   #doesnt work on labelled variables that were imported via haven package

# install.packages('labelled')
require(labelled)  # to be able to use the recode function on haven labelled variables

# inspect variable
str(cv08_haven$var006n)

attr(cv08_haven$var006n, "labels")

table(cv08_haven$var006n, useNA = "always")

# recode values, all missings as one value
cv08_haven <- mutate(cv08_haven, educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2,
    `4` = 2, `5` = 2, `6` = 3, `7` = 3, `8` = 3, `10` = -9), .keep_value_labels = FALSE)

# replace missing values with NA.
cv08_haven <- mutate(cv08_haven, educ3 = na_if(educ3, -9))

# make educ3 a factor
cv08_haven <- mutate(cv08_haven, educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary",
    "tertiary")))

# check
table(cv08_haven$educ3, useNA = "always")

#the piping operator 
  # %>%, 
  # the left is transported to the (first argument of the) function on the right. Thus, in the example below, you see that in the second and third call to mutate I don’t have to tell the function which dataset I am using.
# in Base R (since 4.1.0) operator: |>

cv08_haven <- mutate(cv08_haven, educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2,
    `4` = 2, `5` = 2, `6` = 3, `7` = 3, `8` = 3, `10` = -9), .keep_value_labels = FALSE) %>%
    mutate(educ3 = na_if(educ3, -9)) %>%
    mutate(educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))

# even more tidy:
cv08_haven <- cv08_haven %>%
    mutate(educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2, `4` = 2, `5` = 2,
        `6` = 3, `7` = 3, `8` = 3, `10` = -9), .keep_value_labels = FALSE) %>%
    mutate(educ3 = na_if(educ3, -9)) %>%
    mutate(educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))

# the MOST tidy:
cv08_haven <- cv08_haven %>%
    mutate(educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2, `4` = 2, `5` = 2,
        `6` = 3, `7` = 3, `8` = 3, `10` = -9, .keep_value_labels = FALSE), educ3 = na_if(educ3, -9),
        educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))
```
----
```{r A.9 Means and counting specific values}
# A.9.1. R Base

# Step 1: Look at the various 
summary(cv08$int055)
summary(cv08$int056)
summary(cv08$int057)

# Step 2: Make numeric 
cv08$int055n <- as.numeric(cv08$int055)
table(cv08$int055n, useNA = "always")

cv08$int056n <- as.numeric(cv08$int056)
table(cv08$int056n, useNA = "always")

cv08$int057n <- as.numeric(cv08$int057)
table(cv08$int057n, useNA = "always")

# Step 3: Define missings and recode
cv08$int055n[cv08$int055n < 5] <- NA
cv08$int055n <- cv08$int055n - 4

cv08$int056n[cv08$int056n < 5] <- NA
cv08$int056n <- cv08$int056n - 4

cv08$int057n[cv08$int057n < 5] <- NA
cv08$int057n <- cv08$int057n - 4

# Step 4: Calculate means
# How does the function mean work in R?
mean(cv08$int055n)  # whoops
mean(cv08$int055n, na.rm = TRUE)  # works.  but not what we want. 
# na.rm -> tells function whether or not to remove NA values from the calculation

mean(c(cv08$int055n, cv08$int056n, cv08$int057n), na.rm = T)  # works but not what we want.

# calculate a mean for each row/respondent. 
testmeans <- rowMeans(cbind(cv08$int055n, cv08$int056n, cv08$int057n), na.rm = T)
head(testmeans)  # yes!
# head() function : Return the First or Last Part of an Object

# What we really want is a mean but only if there is a maximum of 1 NA in the three variables.
# lets first count how many missings we have for each respondent
nmis <- rowSums(is.na(cbind(cv08$int055n, cv08$int056n, cv08$int057n)))

# ?is.na ?rowSums
testmeans <- ifelse(nmis < 2, testmeans, NA)

# add the calculated means to our dataset
cv08$int_mean <- testmeans

----

----

# References

>>>>

---
title: "week3 - class RSelenium"
author: "Verooo"
date: "2024-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

# see: https://jochemtolsma.github.io/webscraping/12scholars.html#13_necessary_packages
```{r}
# start with clean workspace
rm(list = ls())

# install.packages("RSelenium")
library(RSelenium)

# Web Driving with R Silenium

```
# citing openalexR in paper!
If you use openalexR in research, please cite: (Priem, Piwowar, and Orr 2022)
